%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% EKKO paper: front matter and preamble filled per ACM template;
%% body is in opera_body.tex; bibliography is references.bib.
%%
%% The first command in your LaTeX source must be the \documentclass
%% command.
%%
\PassOptionsToPackage{table}{xcolor}
\documentclass[sigconf]{acmart}
%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

%% Extra packages required by opera_body.tex (acmart already loads amsmath and booktabs; do not load amssymb---conflicts with newtxmath/\Bbbk).
\usepackage{multirow}
\usepackage{bbm}
\usepackage{bm}
\usepackage{array}
\usepackage{diagbox}
\usepackage{bbding}
\usepackage{colortbl}
\usepackage{tabularx}
\usepackage{xspace}
\usepackage{pifont}

\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

\newcommand{\ie}{\textit{i.e.}\xspace}
\newcommand{\eg}{\textit{e.g.}\xspace}
\newcommand{\etc}{\textit{etc.}\xspace}
\newcommand{\Fref}[1]{Figure~\ref{#1}}
\newcommand{\Eref}[1]{Equation~(\ref{#1})}
\newcommand{\eref}[1]{\Eref{#1}}
\newcommand{\Tref}[1]{Table~\ref{#1}}
\newcommand{\Sref}[1]{Section~\ref{#1}}
\newcommand{\sref}[1]{\Sref{#1}}

\graphicspath{{figures/}}

%% Rights and conference metadata (replace with values from your rights form when final).
\setcopyright{none}
\copyrightyear{2026}
\acmYear{2026}
\acmDOI{10.1145/XXXXXXX.XXXXXXX}
\acmConference[MM '26]{Proceedings of ACM Multimedia 2026}{TBD}{TBD}
\acmISBN{978-1-4503-XXXX-X/2026/XX}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

\title[EKKO]{EKKO: Mitigating Hallucinations in Multimodal Large Language Models via Evaluative Knowledge-Kernel Oracles}

\author{Kaiwen Deng}
\email{dkw24@mails.tsinghua.edu.cn}
\affiliation{%
  \institution{Tsinghua Shenzhen International Graduate School, Tsinghua University}
  \city{Shenzhen}
  \country{China}}

\author{Wenming Yang}
\authornote{Corresponding author.}
\email{yang.wenming@sz.tsinghua.edu.cn}
\affiliation{%
  \institution{Shenzhen International Graduate School \& Department of Electronic Engineering, Tsinghua University}
  \city{Shenzhen}
  \country{China}}

\renewcommand{\shortauthors}{Deng and Yang}

\begin{abstract}
Multimodal Large Language Models (MLLMs) often hallucinate because autoregressive decoding can favor locally plausible tokens that are insufficiently grounded in visual evidence. Existing decode-time mitigation methods mainly address historical failures through retrospective penalties, but they do not explicitly evaluate whether a candidate token is supported by query-relevant visual regions and whether its short-horizon continuation will remain visually grounded. In this paper, we propose EKKO, a novel inference-time decoding framework with Evaluative Knowledge-Kernel Oracles, for mitigating hallucinations in vision-language generation. EKKO evaluates candidate tokens through three coordinated signals: a past signal that preserves rollback-based protection against historical textual inertia, a current signal that measures query-conditioned visual support via anchor-aware region weighting, and a future signal that probes short-horizon continuation tendency through prospective rollout. This design enables EKKO to suppress risky candidates that appear locally plausible at the current step but are likely to trigger text-dominated continuation. To keep the method practical, EKKO performs anchor generation before autoregressive decoding and uses shared-KV batched rollout during candidate evaluation. Experiments on multiple MLLM benchmarks, including POPE, CHAIR, and MMBench, demonstrate that EKKO improves visual grounding and reduces hallucination across diverse multimodal generation settings.

% Konwledge-Kernel Anchor：对应看现在（present）的区域视觉授权
% Evaluative Oracle Rollout：想将来（Future）的轨迹沙盘推演

% Hallucination, posed as a pervasive challenge of multi-modal large language models (MLLMs), has significantly impeded their real-world usage that demands precise judgment.
% Existing methods mitigate this issue with either training with specific designed data or inferencing with external knowledge from other sources, incurring inevitable additional costs.
% In this paper, we present \textbf{OPERA}, a novel MLLM decoding method grounded in an \textbf{O}ver-trust \textbf{Pe}nalty and a \textbf{R}etrospection-\textbf{A}llocation strategy, serving as a nearly \textbf{free lunch} to alleviate the hallucination issue without additional data, knowledge, or training.
% Our approach begins with an interesting observation that, most hallucinations are closely tied to the knowledge aggregation patterns manifested in the self-attention matrix, \ie, MLLMs tend to generate new tokens by focusing on a few summary tokens, but not all the previous tokens.
% Such partial over-trust inclination results in the neglecting of image tokens and describes the image content with hallucination.
% Based on the observation, OPERA introduces a penalty term on the model logits during the beam-search decoding to mitigate the over-trust issue, along with a rollback strategy that retrospects the presence of summary tokens in the previously generated tokens, and re-allocate the token selection if necessary.
% With extensive experiments, OPERA shows significant hallucination-mitigating performance on different MLLMs and metrics, proving its effectiveness and generality.
% Our code is available at: \url{https://github.com/shikiw/OPERA}.
\end{abstract}

\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010147.10010257</concept_id>
  <concept_desc>Computing methodologies~Machine learning</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010405.10010444</concept_id>
  <concept_desc>Applied computing~Computer vision</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010147.10010178.10010179</concept_id>
  <concept_desc>Computing methodologies~Natural language processing</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computing methodologies~Machine learning}
\ccsdesc[300]{Applied computing~Computer vision}
\ccsdesc[300]{Computing methodologies~Natural language processing}

\keywords{multimodal large language models, hallucination mitigation, decoding, beam search, self-attention, OPERA}

%%
%% Figures omitted in opera_body.tex (stub labels only) to avoid float/bibliography interleaving.
%%
\maketitle

\input{opera_body}

\begin{acks}
This work is supported in part by the Natural Science Foundation of China under Grant 62121002, U20B2047, U2336206, 62372423, and 62102386.
This work is partially supported by the National Key R\&D Program of China (2022ZD0160201), and Shanghai Artificial Intelligence Laboratory.
\end{acks}

\clearpage
\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\end{document}
